\bf Model Visualization Techniques for a Social Network Model

### 2017-02-20 paper.Rnw
change title
coulds and shoulds go out

\if0\blind
{
  \title{\bf Visual Diagnsotics of SAO Models}
  \author{Samantha Tyner\thanks{
    The authors gratefully acknowledge funding from the National Science Foundation Grant \# DMS 1007697. All data collection has been conducted with approval from the Institutional Review Board IRB 10-347}\hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visual Diagnsotics of SAO Models}
\end{center}
  \medskip
} \fi

@@

The guiding principles of model visualization are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that could benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization will provide researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

>>>>

\if0\blind
{
  \title{\bf Model Visualization Techniques for a Social Network Model}
  \author{Samantha Tyner\thanks{
    The authors gratefully acknowledge funding from the National Science Foundation Grant \# DMS 1007697. All data collection has been conducted with approval from the Institutional Review Board IRB 10-347}\hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Model Visualization Techniques for a Social Network Model}
\end{center}
  \medskip
} \fi

@@

The guiding principles XXX cite XXX of model visualization are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that  benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization  provides researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robbins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.
